Elephant Seal Age JAE Intraspecific Variation

### 2023-12-06 paper/paperv3.qmd
fixing pup sex figure

---
title: "Elephant Seal Age JAE Intraspecific Variation"
format: 
  html: 
    echo: false
    fig-width: 4.5
    fig-height: 4.5
    code-fold: true
    code-tools: true
    warning: false
    eval: false
editor: visual
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false  
#| eval: true
library(ggthemes)
library(cowplot)
library(lme4)
library(lmerTest) # Loading this library will automatically append p-values to lmer models estimated using something called Satterthwaite's method
library(merTools)
library(tidyverse)
library(ggeffects)
library(patchwork)

@@

```{r}
#| label: age_breeding_mod
#| eval: true

@@

#p values are also rounded....

@@

breed_mod_summary$coefficients %>% 
  as_tibble() %>% 
  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
         across(c(Estimate, `Std. Error`, `z value`), 
                \(x) format_signif(x, digits = 2)),
         `Pr(>|z|)` = format_pval(`Pr(>|z|)`, digits = 3)) %>%
  relocate(Parameter) %>% 
  knitr::kable()
```

@@

surv_recruit_summary %>% 
  ggplot(aes(x = age, y = mean, color = metric, fill = metric)) + 
  geom_ribbon(aes(ymin = mean_lwr, ymax = mean_upr, alpha = metric),
              surv_recruit_pred,
              color = NA) +
  geom_pointrange(aes(ymin = low, ymax = high)) +
  geom_line(aes(linetype = metric),
            surv_recruit_pred) +
  scale_y_continuous(labels = scales::percent) + 
  scale_color_manual(values = c("#829BD1", "#E9B668")) + 
  scale_fill_manual(values = c("#829BD1", "#E9B668")) + 
  scale_linetype_manual(values = c("solid", "dashed")) +
  scale_alpha_manual(values = c(0.25, 0.1)) +
  expand_limits(y = c(0, 0.5)) +
  labs(x = "Maternal Age (Years)", y = "% Pups (Mean ± SE)") +
  theme_few() + 
  theme(legend.position = c(0.1,0.97), 
        legend.justification = c(0,1), 
        legend.title = element_blank(), 
        legend.direction = "horizontal", 
        axis.title = element_text(size = 13))

#Why is recruitment higher than survival in some points? Add the quadratic model? Reverse order of legend in the figure
```

@@

pup_sex <- sealdat %>% 
  filter(pupsex %in% c("M", "F"), 
         age < 19) %>% 
  group_by(age, pupsex) %>%
  count(pupsex) %>% 
  pivot_wider(values_from = n, names_from = pupsex) %>% 
  mutate(num = M + F,
         percf = F / num,
         se = (percf * (1 - percf) / num)^0.5) %>% 
  ggplot(aes(x = age, y = percf, weight = num)) +
  geom_hline(yintercept = 0.5, lty = 2, color = "#829BD1") +
  geom_point() +
  geom_errorbar(aes(ymin = percf - se, ymax = percf + se), width = 0) +
  scale_y_continuous(labels = scales::percent, limits = c(0, NA)) +
  labs(x = "Maternal Age (Years)", y = "% Male Pups (Mean ± SE)") +
  theme_few() + 
  theme(axis.title = element_text(size = 13))

pup_sex

@@

Results of the model for pup sex ratio:

```{r}
#probably will need to take out animalID
pup_sex_mod <- glmer(as.factor(pupsex) ~ age + (1 | animalID) + (1 | year),
                     data = filter(sealdat, 
                                   pupsex %in% c("M", "F"), 
                                   age < 19), 
                     family = "binomial")
summary(pup_sex_mod)

#quadratic

pup_sex_quad <- glmer(as.factor(pupsex) ~ poly(age, 2) + (1 | animalID) + (1 | year),
                            data = filter(sealdat, 
                                   pupsex %in% c("M", "F"), 
                                   age < 19),
                            family = "binomial")
summary(pup_sex_quad)

@@

>>>>

---
title: "Elephant Seal Age JAE Intraspecific Variation"
format: 
  html: 
    echo: false
    fig-width: 4.5
    fig-height: 4.5
    code-fold: true
    code-tools: true
    warning: false
editor: visual
bibliography: references.bib
---

```{r}
#| label: setup
#| include: false  
library(ggthemes)
library(cowplot)
library(lme4)
library(lmerTest) # Loading this library will automatically append p-values to lmer models estimated using something called Satterthwaite's method
library(merTools)
library(tidyverse)
library(ggeffects)
library(patchwork)

@@

```{r}
#| label: age_breeding_mod

@@

#change rightmost value to say p value 
#add an extra sig fig to all columns
#p values are also rounded....

@@

breed_mod_summary$coefficients %>% 
  as_tibble() %>% 
  mutate(Parameter = c("Intercept", "Age (linear)", "Age (quadratic)"),
         across(c(Estimate, `Std. Error`, `z value`), 
                \(x) format_signif(x, digits = 3)),
         P_value = format_pval(`Pr(>|z|)`, digits = 2)) %>%
  select(-`Pr(>|z|)`) %>% 
  relocate(Parameter) %>% 
  knitr::kable()
```

@@

surv_recruit_summary %>% 
  ggplot(aes(x = age, y = mean, color = metric, fill = metric)) + 
  geom_ribbon(aes(ymin = mean_lwr, ymax = mean_upr, alpha = metric),
              surv_recruit_pred,
              color = NA) +
  geom_pointrange(aes(ymin = low, ymax = high)) +
  geom_line(aes(linetype = metric),
            surv_recruit_pred) +
  scale_y_continuous(labels = scales::percent) + 
  scale_color_manual(values = c("#829BD1", "#E9B668")) + 
  scale_fill_manual(values = c("#829BD1", "#E9B668")) + 
  scale_linetype_manual(values = c("solid", "dashed")) +
  scale_alpha_manual(values = c(0.25, 0.1)) +
  expand_limits(y = c(0, 0.5)) +
  labs(x = "Maternal Age (Years)", y = "Pup Survival or Recruitment/n") +
  theme_few() + 
  theme(legend.position = c(0.1,0.97), 
        legend.justification = c(0,1), 
        legend.title = element_blank(), 
        legend.direction = "horizontal", 
        axis.title = element_text(size = 13))

#Why is recruitment higher than survival in some points? Add the quadratic model? Reverse order of legend in the figure
#caption - error bars show SE. Points are mean values for maternal age. Make sure that language in the caption matches both the model and the data 
```

@@

pup_sex_data <- sealdat %>% 
  filter(pupsex %in% c("M", "F"), 
         age < 19) %>% 
  mutate(is_male = ifelse(pupsex == "M", 1, 0))

#linear
pup_sex_mod <- glmer(is_male ~ age + (1 | animalID) + (1 | year),
                     data = pup_sex_data, 
                     family = "binomial")
summary(pup_sex_mod)

# We fit a quadratic equation, but only the linear term was significant, so the linear model is the one that is plotted.

pup_sex_pred <- ggpredict(pup_sex_mod, terms = c("age [all]")) %>% 
  as_tibble() %>% 
  transmute(age = x, 
            percm = predicted, 
            percm_lwr = conf.low, 
            percm_upr = conf.high)

pup_sex_data %>% 
  group_by(age, pupsex) %>%
  summarize(n = n(), .groups = "drop") %>% 
  pivot_wider(values_from = n, names_from = pupsex) %>% 
  mutate(num = M + F,
         percm = M / num,
         se = (percm * (1 - percm) / num)^0.5) %>% 
  ggplot(aes(x = age, y = percm)) +
  geom_hline(yintercept = 0.5, color = "gray75", linetype = "dashed") +
  geom_ribbon(aes(ymin = percm_lwr, ymax = percm_upr), 
              pup_sex_pred,
              fill = "#829BD1",
              alpha = 0.25) + 
  geom_line(data = pup_sex_pred, 
            color = "#829BD1", 
            linewidth = 1) + 
  geom_pointrange(aes(ymin = percm - se, ymax = percm + se), 
                  size = 0.25) +
  scale_y_continuous(labels = scales::percent, limits = c(0, 1)) +
  labs(x = "Maternal Age (Years)", y = "% Male Pups (Mean ± SE)") +
  theme_few() + 
  theme(axis.title = element_text(size = 13))

@@

@@

#average the averages - global average of age-specific averages 
```
